---
title: "diffbind_step8_assign_peaks_enhancers"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Goal
Use loop info from et al from P22 to increase the number of mapped peaks.

# Strategy
1. Get all loops where one of the bins from the pair overlap a promoter region
2. Define the overlapping bin as the "promoter bin" and the other one as the "distal bin"
3. Overlap each no-promoter Zic peak to "distal bins"
4. Map the "Zic bound distal bin" to the gene data from its "promoter bin" pair.
5. Combine all the data to look something like this | distal | promoter | intersected zic peaks | gene info |

# Read data 

Read annotations
```{r}
peaks_all_anno <- readRDS(file = "../data_output/diffbind_cutnrun_zic/peaks_7v1_all_annotation.Rds")

```

Read bin pairs
```{r}
bin_pairs <- read_tsv("../data_input/bedpe/merged_loops.bedpe",comment = "# ", guess_max = 10000)
bin_pairs$loop_id <- paste("loop",seq(1,nrow(bin_pairs)),sep = "-")
bin_pairs$`#chr1` <- paste0("chr",bin_pairs$`#chr1`)
bin_pairs$chr2 <- paste0("chr",bin_pairs$chr2)

bin_pairs_left <- GRanges(bin_pairs$`#chr1`,
        IRanges(bin_pairs$x1,
                bin_pairs$x2),
        loop_id=bin_pairs$loop_id)

bin_pairs_right <- GRanges(bin_pairs$chr2,
        IRanges(bin_pairs$y1,
                bin_pairs$y2),
        loop_id=bin_pairs$loop_id)

```

Read promoters
```{r}
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
mm10_tss <- promoters(genes(TxDb.Mmusculus.UCSC.mm10.knownGene, columns = "gene_id"),upstream =1, downstream=0)
```

# 1. Get all loops where one of the bins from the pair overlap a promoter region

Note: I am selecting only those bins that overlap a single tss
```{r}
bin_pairs_left_olaps <- findOverlaps(mm10_tss, bin_pairs_left, type = "within")
bin_pairs_right_olaps <- findOverlaps(mm10_tss, bin_pairs_right, type = "within")

# Keep bins overlap a single TSS 
bin_pairs_left_olaps <- bin_pairs_left_olaps[! (duplicated(bin_pairs_left_olaps@to) | duplicated(bin_pairs_left_olaps@to, fromLast = TRUE))]

# Keep bins overlap a single TSS 
bin_pairs_right_olaps <- bin_pairs_right_olaps[! (duplicated(bin_pairs_right_olaps@to) | duplicated(bin_pairs_right_olaps@to, fromLast = TRUE))]


```

# 2. Define the overlapping bin as the "promoter bin" and the other one as the "distal bin"

```{r}


bin_pairs_left_tss <- bin_pairs_left[bin_pairs_left_olaps@to]
bin_pairs_left_tss$gene_id <- mm10_tss[bin_pairs_left_olaps@from]$gene_id

bin_pairs_right_tss <- bin_pairs_right[bin_pairs_right_olaps@to]
bin_pairs_right_tss$gene_id <- mm10_tss[bin_pairs_right_olaps@from]$gene_id

# Remove loops where there is a promoter in both sides of loop 
right_both <- bin_pairs_right_tss$loop_id %in% bin_pairs_left_tss$loop_id
left_both <- bin_pairs_left_tss$loop_id %in% bin_pairs_right_tss$loop_id

bin_pairs_right_tss <- bin_pairs_right_tss[!right_both]
bin_pairs_left_tss <- bin_pairs_left_tss[!left_both]

promoters_distal_df <- rbind(left_join(as.data.frame(bin_pairs_right_tss), as.data.frame(bin_pairs_left), by = "loop_id"),
left_join(as.data.frame(bin_pairs_left_tss), as.data.frame(bin_pairs_right), by = "loop_id"))

colnames(promoters_distal_df) <- str_replace(colnames(promoters_distal_df),".x","_promoter_bin")

colnames(promoters_distal_df) <- str_replace(colnames(promoters_distal_df),".y","_distal_bin")


distal_bins <- makeGRangesFromDataFrame(promoters_distal_df,keep.extra.columns = T,
                         seqnames.field = "seqnames_distal_bin",
                         start.field = "start_distal_bin",
                         end.field = "end_distal_bin")
```

# 3. Overlap each no-promoter Zic peak to "distal bins"
4. Map the "Zic bound distal bin" to the gene data from its "promoter bin" pair.

```{r}
peaks_all_anno@anno$peak_id <- seq(1,length(peaks_all_anno@anno))

zic_peaks_non_promoter <- peaks_all_anno@anno[!peaks_all_anno@detailGenomicAnnotation$Promoter]
```

```{r}
# Get peaks midpoints 
zic_peaks_non_promoter_midpoints <- resize(zic_peaks_non_promoter, width = 1, fix = "center")


peaks_distal_olaps <- findOverlaps(zic_peaks_non_promoter_midpoints, distal_bins, type = "within")


# Annotate peaks with their loop id
peaks_all_anno@anno$loop_id <- NA 

peaks_all_anno@anno$loop_id[peaks_all_anno@anno$peak_id[peaks_distal_olaps@from]] <- distal_bins$loop_id[peaks_distal_olaps@to]

```


5. Combine all the data to look something like this | distal | promoter | intersected zic peaks | gene info |

```{r}
peaks_distal_annotated <- as.data.frame(peaks_all_anno@anno)


peaks_distal_annotated <- dplyr::left_join(peaks_distal_annotated, promoters_distal_df, by = "loop_id")

```


Write to table all peaks annotated

```{r}
write_tsv(peaks_distal_annotated,file = "../data_output/diffbind_cutnrun_zic/distal_p22_annotation_all_peaks.tsv")
```




