---
title: "diffbind_step8_assign_peaks_enhancers"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Goal
Use enhancer - promoter pairs from et al from P4 to increase the number of mapped peaks.

# Strategy
1. create a bed file of the enhancers and promoters separately while having a unique identifier for the pair ( e.g.  chr1 5020841 5020906 loop1 (for the enhancer) and chr1 5069785 5070785 loop_1 for the promoter)
2. use bedtools intersect to find the intersecting Zic peaks ( in your case you can read this into R and use GRanges to intersect them)
3. Map the promoter data to the nearest gene. Since this is  promoter data, I expect that the annotated promoters should not be far from its nearest gene.
4. Combine all the data to look something like this | enhancer | promoter | intersected zic peaks | nearest gene to promoter|

# Read annotations
```{r}
peaks_7v1_anno <- readRDS("../data_output/diffbind_cutnrun_zic/peaks_7v1_up_annotation.Rds")
peaks_1v7_anno <- readRDS(file = "../data_output/diffbind_cutnrun_zic/peaks_7v1_down_annotation.Rds")
peaks_all_anno <- readRDS(file = "../data_output/diffbind_cutnrun_zic/peaks_7v1_all_annotation.Rds")

```

# Read enhancer-promoter pairs
```{r}
promoter_enhancer_pairs <- readxl::read_xlsx("../data_input/bedpe/GSE164360_Hi-C_Summary.xlsx",guess_max = 10000)
promoter_enhancer_pairs$loop_id <- paste("loop",seq(1,nrow(promoter_enhancer_pairs)),sep = "-")
```


# Create granges for enhancers and promoters separately

create a bed file of the enhancers and promoters separately while having a unique identifier for the pair ( e.g.  chr1 5020841 5020906 loop1 (for the enhancer) and chr1 5069785 5070785 loop_1 for the promoter)

```{r}
library(GenomicRanges)
pairs_enhancers <- makeGRangesFromDataFrame(promoter_enhancer_pairs, keep.extra.columns = T,
                         seqnames.field = "enhancer_chr",
                         start.field = "enhancer_start",
                         end.field = "enhancer_end")

pairs_promoters <- makeGRangesFromDataFrame(promoter_enhancer_pairs, keep.extra.columns = T,
                         seqnames.field = "promoter_chr",
                         start.field = "promoter_start",
                         end.field = "promoter_end",
                         strand.field = "gene_strand")


```

# Intersect enhancers to peaks
Note: For multiple peaks there is more than one enhancer that overlaps a peak  
```{r}
# Find overlaps
peaks_enhancers_olaps <- findOverlaps(peaks_all_anno@anno, pairs_enhancers)
# For those peaks that overlap more than one enhancer, select the enhancer closest to the peak mid-point
peak2enhancer_id_list <- split(peaks_enhancers_olaps@to, peaks_enhancers_olaps@from)
# Get peaks midpoints 
peaks_resized <- resize(peaks_all_anno@anno, width = 1, fix = "center")

for (peak in names(peak2enhancer_id_list)){
  # If a peak overlaps more than one enhancer
  if(length(peak2enhancer_id_list[[peak]])>1){
    # Get the indices of overlapping enhancers
    enhancers <- peak2enhancer_id_list[[peak]]
    # Get the distance between peak and nearest enhancers 
    distonear <- distanceToNearest(peaks_all_anno@anno[as.numeric(peak)],
                                  pairs_enhancers[enhancers])
    # Change enhancer to the nearest one, in extreme case where there are still ties, select the first one just to guarantee a single assignment
    peak2enhancer_id_list[[peak]] <- enhancers[distonear@to[1]]
    
  }
}
# Annotate peaks with their loop id
peaks_all_anno@anno$loop_id <- NA 

peaks_all_anno@anno$loop_id[as.numeric(names(peak2enhancer_id_list))] <- pairs_enhancers$loop_id[unlist(peak2enhancer_id_list)]

```


# Intersect promoters to genes

Note: for the first pass, even though we are doing these promoter / gene overlaps, in the final table we'll use the genes they assigned 
```{r}
pairs_promoters_assigned <- pairs_promoters[pairs_promoters$loop_id %in% pairs_enhancers$loop_id[unlist(peak2enhancer_id_list)]]

library(TxDb.Mmusculus.UCSC.mm10.knownGene)
mm10_tss <- promoters(genes(TxDb.Mmusculus.UCSC.mm10.knownGene, columns = "gene_id"),upstream =500, downstream=501)


distonear_promoters <- distanceToNearest(pairs_promoters_assigned,mm10_tss)

# There are no duplicates but to ensure 1-1 mappings, arbitrarily select first "nearest" one
distonear_promoters <- distonear_promoters[!duplicated(distonear_promoters@from)]

pairs_promoters_assigned$gene_new_assignment <- mm10_tss[distonear_promoters@to]$gene_id
pairs_promoters_assigned$gene_new_assignment_distance <- distonear_promoters@elementMetadata$distance 

library(clusterProfiler)
library(org.Mm.eg.db)


gene_id_mappings <- bitr(pairs_promoters_assigned$gene_new_assignment,fromType = "ENTREZID",toType = "SYMBOL",OrgDb = org.Mm.eg.db)

rownames(gene_id_mappings) <- gene_id_mappings$ENTREZID

pairs_promoters_assigned$gene_new_assignment_symbol <- gene_id_mappings[pairs_promoters_assigned$gene_new_assignment,"SYMBOL"]

loop_gene_assignments <- as.data.frame(elementMetadata(pairs_promoters_assigned)[,c("loop_id","gene_new_assignment","gene_new_assignment_symbol","gene_new_assignment_distance")])

```

# Combine data

4. Combine all the data to look something like this | enhancer | promoter | intersected zic peaks | nearest gene to promoter|

```{r}
peaks_enhancer_annotated <- as.data.frame(peaks_all_anno@anno)


peaks_enhancer_annotated <- dplyr::left_join(peaks_enhancer_annotated, promoter_enhancer_pairs, by = "loop_id")

peaks_enhancer_annotated <- dplyr::left_join(peaks_enhancer_annotated, loop_gene_assignments, by = "loop_id")

```

All peaks annotated

```{r}
write_tsv(peaks_enhancer_annotated,file = "../data_output/diffbind_cutnrun_zic/enhancer_annotation_all_peaks.tsv")
```


All annotated loops 
```{r}
loops_annotated_all <- peaks_enhancer_annotated %>%
  drop_na(loop_id) 
write_tsv(loops_annotated_all,file = "../data_output/diffbind_cutnrun_zic/enhancer_annotation_all_loops.tsv")

```

Annotated loops, relevant columns (enhancers, promoters, peaks, gene)
```{r}
loops_annotated <- loops_annotated_all%>%
  select(
         enhancer_chr, enhancer_start, enhancer_end,
         promoter_chr, promoter_start, promoter_end,
         gene_strand,
         seqnames, start, end, strand, baseMean, log2FoldChange, padj,
         GeneID)


write_tsv(loops_annotated,file = "../data_output/diffbind_cutnrun_zic/enhancer_annotation_loops.tsv")

```

D1 table
```{r}
loops_annotated_d1 <- loops_annotated %>%
  filter(padj < 0.05 & log2FoldChange < 0)

write_tsv(loops_annotated_d1,file = "../data_output/diffbind_cutnrun_zic/enhancer_annotation_loops_d1.tsv")

```


D7 table
```{r}
loops_annotated_d7 <- loops_annotated %>%
  filter(padj < 0.05 & log2FoldChange > 0)

write_tsv(loops_annotated_d7,file = "../data_output/diffbind_cutnrun_zic/enhancer_annotation_loops_d7.tsv")

```






