---
title: "diffbind_step3_dbplots"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Read results

```{r}
res_3v1 <- readRDS("../data_output/diffbind_cutnrun_zic/res_3v1_ashr.Rds")
res_5v3 <- readRDS("../data_output/diffbind_cutnrun_zic/res_5v3_ashr.Rds")
res_7v5 <- readRDS("../data_output/diffbind_cutnrun_zic/res_7v5_ashr.Rds")
res_7v1 <- readRDS("../data_output/diffbind_cutnrun_zic/res_7v1_ashr.Rds")
res_all <- readRDS("../data_output/diffbind_cutnrun_zic/res_all_ashr.Rds")
```

# Plots
```{r}
library(tidyverse)
```

## P value histogram

```{r}
res_3v1 %>%
  as_tibble() %>%
  ggplot(aes(pvalue)) +
  geom_histogram()

res_5v3 %>%
  as_tibble() %>%
  ggplot(aes(pvalue)) +
  geom_histogram()

res_7v5 %>%
  as_tibble() %>%
  ggplot(aes(pvalue)) +
  geom_histogram()

res_7v1 %>%
  as_tibble() %>%
  ggplot(aes(pvalue)) +
  geom_histogram()

res_all %>%
  as_tibble() %>%
  ggplot(aes(pvalue)) +
  geom_histogram()

```

## MA 

```{r}
plotMA(res_3v1, main = "DIV3-DIV1", alpha = 0.05, ylim = c(-3,3)) 
plotMA(res_5v3, main = "DIV5-DIV3", alpha = 0.05, ylim = c(-1,1))
plotMA(res_7v5, main = "DIV7-DIV5", alpha = 0.05, ylim = c(-1,1))
plotMA(res_7v1, main = "DIV7-DIV1", alpha = 0.05, ylim = c(-3,3))

```

## Volcano

```{r}
plot_volcano <- function(res){
  res %>%
    as_tibble() %>%
    ggplot(aes(log2FoldChange, -10 * log10(padj))) +
    geom_point(aes(color = padj < 0.05 )) +
    scale_color_manual(values = c("FALSE" = "gray","TRUE"="blue")) +
    theme_classic() +
    theme(aspect.ratio = 1) 
}
```

```{r}
plot_volcano(res_3v1) + ggtitle("DIV3-DIV1")
plot_volcano(res_5v3) + ggtitle("DIV5-DIV3")
plot_volcano(res_7v5) + ggtitle("DIV7-DIV5")
plot_volcano(res_7v1) + ggtitle("DIV7-DIV1")

```

## Heatmap

Get vst counts
```{r}
deseq_obj_contrasts <- readRDS("../data_output/diffbind_cutnrun_zic/deseq_obj_contrasts.Rds")

vsd <- vst(deseq_obj_contrasts, blind=FALSE)


```

```{r}
library(ComplexHeatmap)


plot_diffbind_heatmap <- function(vst_tab, res_obj){
  scaled_vst <- t(scale(t(assay(vst_tab[tidyr::replace_na(res_obj$padj,1) < 0.05,]))))
  colnames(scaled_vst) <- paste((str_split(colnames(scaled_vst),"-") %>% map_chr(2)),
                                (str_split(colnames(scaled_vst),"-") %>% map_chr(1)),sep = "-")
  
  scaled_vst <- scaled_vst[, order(colnames(scaled_vst))]
  
  Heatmap(scaled_vst ,
          cluster_columns = FALSE,
          show_row_dend = FALSE,
          col = viridis::inferno(n = 2,begin = 0.2, end = 0.8))
  
} 




```

```{r}
plot_diffbind_heatmap(vsd, res_3v1) + ggtitle("DIV3-DIV1")
plot_diffbind_heatmap(vsd, res_5v3)+ ggtitle("DIV5-DIV3")
plot_diffbind_heatmap(vsd, res_7v5)+ ggtitle("DIV7-DIV5")
plot_diffbind_heatmap(vsd, res_7v1) + ggtitle("DIV7-DIV1")
plot_diffbind_heatmap(vsd, res_all)+ ggtitle("Any timepoint")
```

DIV7-DIV1 K means

```{r}
scaled_vst <- t(scale(t(assay(vsd[tidyr::replace_na(res_7v1$padj,1) < 0.05,]))))
colnames(scaled_vst) <- paste((str_split(colnames(scaled_vst),"-") %>% map_chr(2)),
                              (str_split(colnames(scaled_vst),"-") %>% map_chr(1)),sep = "-")

scaled_vst <- scaled_vst[, order(colnames(scaled_vst))]

Heatmap(scaled_vst ,
        cluster_columns = FALSE,
        show_row_dend = FALSE,
        row_km = 2,
        col = viridis::inferno(n = 2,begin = 0.2, end = 0.8))

```

DIV7-DIV1 K means

```{r}
scaled_vst <- t(scale(t(assay(vsd[tidyr::replace_na(res_7v1$padj,1) < 0.05 & abs(tidyr::replace_na(res_7v1$log2FoldChange,1)) > 1.5,]))))
colnames(scaled_vst) <- paste((str_split(colnames(scaled_vst),"-") %>% map_chr(2)),
                              (str_split(colnames(scaled_vst),"-") %>% map_chr(1)),sep = "-")

scaled_vst <- scaled_vst[, order(colnames(scaled_vst))]

Heatmap(scaled_vst ,
        cluster_columns = FALSE,
        show_row_dend = FALSE,
        row_km = 2,
        col = viridis::inferno(n = 2,begin = 0.2, end = 0.8))

```

# Notes: 

- In general there seems to be three large groups:

- Peaks that loose binding 1 -> 7
- Peaks that gain binding 1 -> 7 
For both of these there might be subgroups of different gain/loss timing (i.e. some binding is gained by day 3, others 5 etc)
- Peaks gain and then loose (gain by 3 then start loosing) Could it be due to cell culture?



