---
title: "diffbind_step10_assign_peaks_loops"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(plyranges)
```

# Goal

Exploration of assignment of in vitro Zic peaks to genes via loops by P4 or P22 cerebellum HiC data

# Read data

Read loop - gene assignments
```{r}
mapped_data <- read_tsv("../data_output/diffbind_cutnrun_zic/melyssa_pipeline/mapped_data.txt")

mapped_data <- drop_na(mapped_data, zic_peak)
colnames(mapped_data) <- paste0("loop_",colnames(mapped_data))

mapped_data <- mapped_data %>%
  separate(loop_zic_peak,sep = ":",into = c("seqnames","startend")) %>%
  separate(startend, sep = "-", into = c("start", "end"))

mapped_data$start <- as.numeric(mapped_data$start) + 1
mapped_data$end <- as.numeric(mapped_data$end)

```


Read diff peak data
```{r}
peaks_all <- readRDS(file = "../data_output/diffbind_cutnrun_zic/peaks_7v1_all_annotation.Rds")

peaks_all_mapped <- as.data.frame(peaks_all@anno)

```

Join tables
```{r}
peaks_all_mapped <- left_join(peaks_all_mapped, mapped_data, by = c("seqnames", "start", "end"))
peaks_all_mapped <- peaks_all_mapped %>%
  distinct()

```

```{r}
peaks_all_mapped %>%
  filter(is.na(loop_id)) 
```

# Assignment P4
```{r}
peaks_all_mapped %>%
  filter(str_detect(loop_id,"young")) %>%
  mutate(
         is_diff = case_when(
           padj < 0.05 & log2FoldChange > 0 ~ "DIV7",
           padj < 0.05 & log2FoldChange < 0 ~ "DIV1",
           is.na(padj) ~ "NA",
           TRUE ~ "ns",
         ),
         is_diff = fct_relevel(is_diff, "DIV1","ns","DIV7")) %>%
  filter(is_diff != "NA") %>%
  select(seqnames, start, end, is_diff) %>%
  distinct() %>%
  group_by(is_diff) %>%
  summarise(npeaks = n()) %>%
  ggplot(aes(x = is_diff, y =npeaks, fill = is_diff, label = npeaks)) +
  geom_col() +
  scale_fill_manual(values = c("ns" = "lightgray","DIV7" = "red","DIV1"="blue"))+
  theme_minimal() +
  theme(aspect.ratio = 1) +
  geom_text(nudge_y = 200)
  
```



# Assignment P22
```{r}
peaks_all_mapped %>%
  filter(str_detect(loop_id,"adult")) %>%
  mutate(
         is_diff = case_when(
           padj < 0.05 & log2FoldChange > 0 ~ "DIV7",
           padj < 0.05 & log2FoldChange < 0 ~ "DIV1",
           is.na(padj) ~ "NA",
           TRUE ~ "ns",
         ),
         is_diff = fct_relevel(is_diff, "DIV1","ns","DIV7")) %>%
  pull(is_diff) %>%
  table()
```


```{r}
peaks_all_mapped %>%
  filter(str_detect(loop_id,"adult")) %>%
  mutate(
         is_diff = case_when(
           padj < 0.05 & log2FoldChange > 0 ~ "DIV7",
           padj < 0.05 & log2FoldChange < 0 ~ "DIV1",
           is.na(padj) ~ "NA",
           TRUE ~ "ns",
         ),
         is_diff = fct_relevel(is_diff, "DIV1","ns","DIV7")) %>%
  filter(is_diff != "NA") %>%
  select(seqnames, start, end, is_diff) %>%
  distinct() %>%
  group_by(is_diff) %>%
  summarise(npeaks = n()) %>%
  ggplot(aes(x = is_diff, y =npeaks, fill = is_diff, label = npeaks)) +
  geom_col() +
  scale_fill_manual(values = c("ns" = "lightgray","DIV7" = "red","DIV1"="blue"))+
  theme_minimal() +
  theme(aspect.ratio = 1) +
  geom_text(nudge_y = 500)
  
```

# Zic peaks overlap

# Read Zic
```{r}
peaks_all_mapped_div7 <- peaks_all_mapped[replace_na(peaks_all_mapped$padj,1) < 0.05 & replace_na(peaks_all_mapped$log2FoldChange,0) > 0,]


peaks_all_mapped_div1 <- peaks_all_mapped[replace_na(peaks_all_mapped$padj ,1) < 0.05 & replace_na(peaks_all_mapped$log2FoldChange,0) < 0,]


peaks_all_mapped_div7_adult <- peaks_all_mapped_div7 %>%
  filter(str_detect(loop_id,"adult")) %>%
  select(seqnames, start, end) %>%
  unique() %>%
  summarise(cts = n()) %>%
  pull(cts)

peaks_all_mapped_div7_young <- peaks_all_mapped_div7 %>%
  filter(str_detect(loop_id,"young")) %>%
  select(seqnames, start, end) %>%
  unique() %>%
  summarise(cts = n()) %>%
  pull(cts)

peaks_all_mapped_div1_adult <- peaks_all_mapped_div1 %>%
  filter(str_detect(loop_id,"adult")) %>%
  select(seqnames, start, end) %>%
  unique() %>%
  summarise(cts = n()) %>%
  pull(cts)

peaks_all_mapped_div1_young <- peaks_all_mapped_div1 %>%
  filter(str_detect(loop_id,"young")) %>%
  select(seqnames, start, end) %>%
  unique() %>%
  summarise(cts = n()) %>%
  pull(cts)

```


```{r}
all_anno <- as.data.frame(peaks_all@anno)
all_anno <- drop_na(all_anno,padj)
all_anno <- drop_na(all_anno,log2FoldChange)
tot_div7 <- sum(all_anno$padj < 0.05 & all_anno$log2FoldChange > 0)
tot_div1 <- sum(all_anno$padj < 0.05 & all_anno$log2FoldChange < 0)

```


# Get overlaps

```{r}

overlaps_df <- data.frame("intersect" =c(peaks_all_mapped_div7_adult,
peaks_all_mapped_div1_adult,
peaks_all_mapped_div7_young,
peaks_all_mapped_div1_young))


overlaps_df$a <- c("peaks_d7", "peaks_d1","peaks_d7","peaks_d1")
overlaps_df$b <- c("p22","p22","p4","p4")

overlaps_df$totals_a <- c(tot_div7,
                          tot_div1,
                          tot_div7,
                          tot_div1)
overlaps_df$percentage_a <- overlaps_df$intersect * 100 / overlaps_df$totals_a


```

```{r}
overlaps_df %>%
  mutate(b = fct_relevel(b,"p4","p22")) %>%
  ggplot(aes(a,percentage_a, fill = a)) +
  geom_col() +
  facet_grid(~b,scales = "free") +
  scale_fill_manual(values = c("peaks_d1" = "#1E88E5",
                               "peaks_d7" = "#D81B60")) +
  theme_light() +
  theme(aspect.ratio = 2) +
  geom_text(aes(label = paste0(round(intersect)," (",round(percentage_a),"%) ")), nudge_y = 1) 

```


Notes:

- I don't understand why there are less assigned loops than overlapping bins?
- I think it might be due to assignment of nearest gene to bins where those bins don't overlap, but then why would the early ones be less?
- I think the P4 loops might not be as close to genes and the P22 are.?  
- Or is it ghe bin size?
- Other than that, it looks like we get a bit more assignments

```{r}
mapped_data[mapped_data$loop_id == "adult_loop_207",]
```


# Basic looping statistics
```{r}
p22_loops <- read_tsv("../data_input/bedpe/merged_loops.bedpe")
p4_loops <- readxl::read_excel("../data_input/bedpe/GSE164360_Hi-C_Summary.xlsx")
p22_loops <- p22_loops[-1,]
```

```{r}
nrow(unique(rbind(as.matrix(p4_loops[,1:3]),
as.matrix(p4_loops[,4:6]))))

```


```{r}

nrow(unique(rbind(as.matrix(p22_loops[,1:3]),
as.matrix(p22_loops[,4:6]))))

```

```{r}
div7_peaks <- readRDS(file = "../data_output/diffbind_cutnrun_zic/peaks_7v1_up_annotation.Rds")
div1_peaks <- readRDS(file = "../data_output/diffbind_cutnrun_zic/peaks_7v1_down_annotation.Rds")

sum(str_detect(div7_peaks@anno$annotation,"Promoter"))
sum(str_detect(div1_peaks@anno$annotation,"Promoter"))


length(unique(div7_peaks@anno[str_detect(div7_peaks@anno$annotation,"Promoter")]$SYMBOL))

length(unique(div1_peaks@anno[str_detect(div1_peaks@anno$annotation,"Promoter")]$SYMBOL))

peaks_all@anno %>%
  as.data.frame() %>%
  drop_na(padj) %>%
  filter(padj > 0.05) %>%
  filter(str_detect(annotation, "Promoter")) %>%
  select(SYMBOL) %>%
  unique()

```

```{r}
mapped_data <- read_tsv("../data_output/diffbind_cutnrun_zic/melyssa_pipeline/mapped_data.txt")

colnames(mapped_data) <- paste0("loop_",colnames(mapped_data))

peaks_all_anno <- as.data.frame(peaks_all@anno)
peaks_all_anno <- peaks_all_anno %>%
  drop_na(padj)


loops <- left_join(as.data.frame(peaks_all@anno), mapped_data)
```
```{r}
peaks_all_mapped %>%
  drop_na(loop_id) %>%
  drop_na(padj) %>%
  drop_na(loop_gene_name) %>%
  filter(padj > 0.05) %>%
  select(seqnames, start, end) %>%
  unique()
```

