---
title: "zic_cutnrun_annotation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Zic CutNRun peaks annotation 

# Load CutNrun and Chip peaks

## load cutnrun peaks as granges objects seacr
```{r}
library(rtracklayer)
library(tidyverse)

cutnrun_seacr_dir <- "../data_output/cutnrun_zic/seacr_peaks/"
cutnrun_seacr_files <- list.files(cutnrun_seacr_dir,pattern = ".bed")
cutnrun_seacr_full <- paste0(cutnrun_seacr_dir,cutnrun_seacr_files)


cutnrun_seacr_granges <- lapply(cutnrun_seacr_full, import.bedGraph)

names(cutnrun_seacr_granges) <- str_split(cutnrun_seacr_files,"_",n=2) %>% map_chr(1)

names(cutnrun_seacr_granges) <- paste0("cutnrun-seacr-",names(cutnrun_seacr_granges))

```

macs
```{r}

cutnrun_macs_dir <- "../data_output/cutnrun_zic/macs_peaks/"
cutnrun_macs_files <- list.files(cutnrun_macs_dir, pattern = "narrowPeak$")
cutnrun_macs_full <- paste0(cutnrun_macs_dir,cutnrun_macs_files)

cutnrun_macs_granges <- lapply(cutnrun_macs_full,
                               import,
                               format = "BED",
                               extraCols = c(singnalValue = "numeric",
                                             pValue = "numeric",
                                             qValue = "numeric",
                                             peak = "integer"))

names(cutnrun_macs_granges) <- str_split(cutnrun_macs_files,"_",n=2) %>% map_chr(1)

names(cutnrun_macs_granges) <- paste0("cutnrun-macs-",names(cutnrun_macs_granges))

```


## load ChIP peaks - previous macs
```{r}
chip_peaks_dir <- "../data_output/bed/"
chip_peaks_files <- list.files(chip_peaks_dir, pattern = "mm10.bed")
chip_peaks_full <- paste0(chip_peaks_dir,chip_peaks_files)
chip_names <- str_split(chip_peaks_files,"_",n=3) %>% map_chr(2)
chip_names <- paste0("ChIP-macs1-","rep",rep(1:2,2),"-",chip_names)

chip_macs1_granges <- lapply(chip_peaks_full, import.bed)

names(chip_macs1_granges) <- chip_names

```

## load ChIP peaks - reprocessed macs
```{r}
chip_macs2_peaks_dir <- "../data_output/chip_zic/chip_zic_macs2_peaks/"
chip_macs2_peaks_files <- list.files(chip_macs2_peaks_dir)
chip_macs2_peaks_files <- chip_macs2_peaks_files[1:4]
chip_macs2_peaks_full <- paste0(chip_macs2_peaks_dir,chip_macs2_peaks_files)
chip_macs2_peaks_names <- c("P7","P7","P60","P60")

chip_macs2_peaks_names <- paste0("ChIP-macs2-","rep",rep(1:2,2),"-",chip_macs2_peaks_names)


chip_macs2_granges <- lapply(chip_macs2_peaks_full,
                               import,
                               format = "BED",
                               extraCols = c(singnalValue = "numeric",
                                             pValue = "numeric",
                                             qValue = "numeric",
                                             peak = "integer"))


names(chip_macs2_granges) <- chip_macs2_peaks_names

```

## load ChIP peaks - reprocessed seacr
```{r}
chip_seacr_peaks_dir <- "../data_output/chip_zic/chip_zic_seacr_peaks/"
chip_seacr_peaks_files <- list.files(chip_seacr_peaks_dir,pattern = "bed$")
chip_seacr_peaks_files <- chip_seacr_peaks_files[1:4]
chip_seacr_peaks_full <- paste0(chip_seacr_peaks_dir,chip_seacr_peaks_files)
chip_seacr_peaks_names <- c("P7","P7","P60","P60")
chip_seacr_peaks_names <- paste0("ChIP-seacr-","rep",rep(1:2,2),"-",chip_seacr_peaks_names)

chip_seacr_granges <- lapply(chip_seacr_peaks_full, import.bedGraph)

names(chip_seacr_granges) <- chip_seacr_peaks_names

```


```{r}
library(ChIPseeker)
library(clusterProfiler)
library(org.Mm.eg.db)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
txdb <- TxDb.Mmusculus.UCSC.mm10.knownGene

```

# Annotate peaks 

```{r}

anno_cutnrun_seacr <- lapply(cutnrun_seacr_granges,
       annotatePeak,
       TxDb = txdb,
             level = "gene",
             assignGenomicAnnotation = T,
             annoDb = "org.Mm.eg.db")

anno_cutnrun_macs <- lapply(cutnrun_macs_granges,
       annotatePeak,
       TxDb = txdb,
             level = "gene",
             assignGenomicAnnotation = T,
             annoDb = "org.Mm.eg.db")

anno_chip_macs1 <- lapply(chip_macs1_granges,
       annotatePeak,
       TxDb = txdb,
             level = "gene",
             assignGenomicAnnotation = T,
             annoDb = "org.Mm.eg.db")


anno_chip_macs2 <- lapply(chip_macs2_granges,
       annotatePeak,
       TxDb = txdb,
             level = "gene",
             assignGenomicAnnotation = T,
             annoDb = "org.Mm.eg.db")


anno_chip_seacr <- lapply(chip_seacr_granges,
       annotatePeak,
       TxDb = txdb,
             level = "gene",
             assignGenomicAnnotation = T,
             annoDb = "org.Mm.eg.db")
```

```{r}
names(anno_cutnrun_seacr) <- paste(str_split(names(anno_cutnrun_seacr),"-",4) %>% map_chr(4),
      str_split(names(anno_cutnrun_seacr),"-",4) %>% map_chr(3),
      sep = "-")

anno_cutnrun_seacr <- anno_cutnrun_seacr[order(names(anno_cutnrun_seacr))]

plotAnnoBar(anno_cutnrun_seacr) + ggtitle("cutnrun-seacr")
```

```{r}
names(anno_cutnrun_macs) <- paste(str_split(names(anno_cutnrun_macs),"-",4) %>% map_chr(4),
      str_split(names(anno_cutnrun_macs),"-",4) %>% map_chr(3),
      sep = "-")

anno_cutnrun_macs <- anno_cutnrun_macs[order(names(anno_cutnrun_macs))]

plotAnnoBar(anno_cutnrun_macs) + ggtitle("cutnrun-macs")

```


```{r}
names(anno_chip_macs1) <- paste(str_split(names(anno_chip_macs1),"-",4) %>% map_chr(4),
      str_split(names(anno_chip_macs1),"-",4) %>% map_chr(3),
      sep = "-")

plotAnnoBar(anno_chip_macs1) + ggtitle("ChIP-macs1")

```

```{r}
names(anno_chip_macs2) <- paste(str_split(names(anno_chip_macs2),"-",4) %>% map_chr(4),
      str_split(names(anno_chip_macs2),"-",4) %>% map_chr(3),
      sep = "-")

plotAnnoBar(anno_chip_macs2) + ggtitle("ChIP-macs2")

```

```{r}
names(anno_chip_seacr) <- paste(str_split(names(anno_chip_seacr),"-",4) %>% map_chr(4),
      str_split(names(anno_chip_seacr),"-",4) %>% map_chr(3),
      sep = "-")

plotAnnoBar(anno_chip_seacr) + ggtitle("ChIP-seacr")

```

Notes: 
- cutnrun - macs2 progressively gains peaks at promoters
- cutnrun - seacr doesn't show this

- for chip, there is inconsistent data between methods, with only seacr showing gain of some peaks at promoters, but don't know if we should read too much into this
