---
title: "zic_cutnrun_chip_overlaps"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Overview

How similar are in vitro and in vivo Zic peaks?

- CutNRun overlap with ChIP peaks
  - Download processed Zic ChIP peaks 
  - Liftover to mm10
  - Load CutNrun and Chip peaks
  - Overlap all vs all
  - Plot overlaps

# Download processed Zic ChIP peaks 

Downloaded from: https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE60731

To: data_input/bed/

# Liftover to mm10
Download liftover file
```{bash}
wget http://hgdownload.soe.ucsc.edu/goldenPath/mm9/liftOver/mm9ToMm10.over.chain.gz
```

Lift files
```{bash}
out=../data_output/bed/
for old_file in ../data_input/bed/*bed;
do
  basename=${old_file%.*}
  basename=${basename##*/}
  new_file=${out}${basename}_mm10.bed
  unmapped=${out}${basename}_mm10_unmapped.bed
  liftOver $old_file ../data_input/liftOver/mm9ToMm10.over.chain $new_file $unmapped
done
```


# Load CutNrun and Chip peaks

## load cutnrun peaks as granges objects
```{r}
library(rtracklayer)
library(tidyverse)

cutnrun_peaks_dir <- "../data_output/cutnrun_zic/seacr_peaks/"
cutnrun_peaks_files <- list.files(cutnrun_peaks_dir,pattern = ".bed")
cutnrun_peaks_full <- paste0(cutnrun_peaks_dir,cutnrun_peaks_files)


cutnrun_peaks_granges <- lapply(cutnrun_peaks_full, import.bedGraph)

names(cutnrun_peaks_granges) <- str_split(cutnrun_peaks_files,"_",n=2) %>% map_chr(1)

names(cutnrun_peaks_granges) <- paste0("cutnrun-",names(cutnrun_peaks_granges))

```

## load ChIP peaks
```{r}
chip_peaks_dir <- "../data_output/bed/chip_zic/"
chip_peaks_files <- list.files(chip_peaks_dir, pattern = "mm10.bed")
chip_peaks_full <- paste0(chip_peaks_dir,chip_peaks_files)
chip_names <- str_split(chip_peaks_files,"_",n=3) %>% map_chr(2)
chip_names <- paste0("ChIP-","rep",rep(1:2,2),"-",chip_names)

chip_peaks_granges <- lapply(chip_peaks_full, import.bed)


names(chip_peaks_granges) <- chip_names


```

# Overlap all vs all

```{r}
library(GenomicRanges)

all_peaks <- GRangesList(c(cutnrun_peaks_granges,chip_peaks_granges))

```



In basepairs

```{r}

all_vs_all_overlap <- combn(1:length(all_peaks),2,
                            FUN = function(x) {
                              intersection <- sum(width(GenomicRanges::intersect(all_peaks[[x[1]]],all_peaks[[x[2]]])))
                              setdiff1 <- sum(width(GenomicRanges::setdiff(all_peaks[[x[1]]],all_peaks[[x[2]]])))
                              setdiff2 <- sum(width(GenomicRanges::setdiff(all_peaks[[x[2]]],all_peaks[[x[1]]])))
                              union <- sum(width(GenomicRanges::union(all_peaks[[x[1]]],all_peaks[[x[2]]])))
                              return(data.frame("intersection" = intersection,
                                                "union" = union,
                                                "set1" = names(all_peaks[x[1]]),
                                                "set2" = names(all_peaks[x[2]]),
                                                "setdiff1" = setdiff1,
                                                "setdiff2" = setdiff2)
                              )
                            },
                            simplify = FALSE)

all_vs_all_overlap <- do.call(rbind,all_vs_all_overlap)


```

By element

```{r}

all_vs_all_overlap_by_element <- combn(1:length(all_peaks),2,
                            FUN = function(x) {
                              intersection <- sum(IRanges::overlapsAny(all_peaks[[x[1]]],all_peaks[[x[2]]]))
                              setdiff1 <- length(all_peaks[[x[1]]]) - intersection
                              setdiff2 <-length(all_peaks[[x[2]]]) - intersection 
                              
                              return(data.frame("intersection" = intersection,
                                                "set1" = names(all_peaks[x[1]]),
                                                "set2" = names(all_peaks[x[2]]),
                                                "setdiff1" = setdiff1,
                                                "setdiff2" = setdiff2)
                              )
                            },
                            simplify = FALSE)

all_vs_all_overlap_by_element <- do.call(rbind,all_vs_all_overlap_by_element)


```

# Plots

```{r}
plot_overlaps <- function(overlaps_df) {
  overlaps_df %>%
  pivot_longer(is.numeric) %>%
    group_by(set1,set2) %>%
    mutate(percentage = value / sum(value)) %>%
    ungroup() %>%
    mutate(name = fct_relevel(name, "setdiff2","setdiff1","intersection")) %>%
    ggplot(aes(y = paste(set1,set2), x = percentage, fill = name)) +
    geom_col(width = 0.5) +
    ylab(NULL) +
    xlab("Percentage Overlap") +
    theme_minimal() 
}
```


## Matched times plots

### Culture day 3 vs in vivo day 7

Basepairs
```{r}
all_vs_all_overlap %>%
  filter( (str_detect(set1, "D3") &
           str_detect(set2, "P7") ) | (
             str_detect(set1, "P7") &
           str_detect(set2, "D3") )
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() 

```

By element
```{r}
all_vs_all_overlap_by_element %>%
  filter( (str_detect(set1, "D3") &
           str_detect(set2, "P7") ) | (
             str_detect(set1, "P7") &
           str_detect(set2, "D3") )
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() 

```
### Culture day 7 vs in vivo day 60

By basepairs
```{r}
all_vs_all_overlap %>%
  filter( (str_detect(set1, "D7") &
           str_detect(set2, "P60") ) | (
             str_detect(set1, "P60") &
           str_detect(set2, "D7") )
           )  %>%
  select(!starts_with("union")) %>%
  plot_overlaps() 
```

By elements
```{r}
all_vs_all_overlap_by_element %>%
  filter( (str_detect(set1, "D7") &
           str_detect(set2, "P60") ) | (
             str_detect(set1, "P60") &
           str_detect(set2, "D7") )
           )  %>%
  plot_overlaps() 
```
### Culture day 1 replicates

By basepair
```{r}
all_vs_all_overlap %>%
  filter( str_detect(set1, "D1") &
           str_detect(set2, "D1") )  %>%
  select(!starts_with("union")) %>%
  plot_overlaps()
```
By element
```{r}
all_vs_all_overlap_by_element %>%
  filter( str_detect(set1, "D1") &
           str_detect(set2, "D1") )  %>%
  select(!starts_with("union")) %>%
  plot_overlaps()
```
### Culture day 3 replicates

By basepair

```{r}
all_vs_all_overlap %>%
  filter( str_detect(set1, "D3") &
           str_detect(set2, "D3") )  %>%
  select(!starts_with("union")) %>%
  plot_overlaps()

```
By elements

```{r}
all_vs_all_overlap_by_element %>%
  filter( str_detect(set1, "D3") &
           str_detect(set2, "D3") )  %>%
  select(!starts_with("union")) %>%
  plot_overlaps()

```
### Culture day 5 replicates

By basepairs
```{r}
all_vs_all_overlap %>%
  filter( str_detect(set1, "D5") &
           str_detect(set2, "D5") )  %>%
  select(!starts_with("union")) %>%
  plot_overlaps()
```
By elements
```{r}
all_vs_all_overlap_by_element %>%
  filter( str_detect(set1, "D5") &
           str_detect(set2, "D5") )  %>%
  select(!starts_with("union")) %>%
  plot_overlaps()
```
### Culture day 7 replicates

```{r}
all_vs_all_overlap %>%
  filter( str_detect(set1, "D7") &
           str_detect(set2, "D7") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps()
```

```{r}
all_vs_all_overlap_by_element %>%
  filter( str_detect(set1, "D7") &
           str_detect(set2, "D7") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps()
```
### Chip P7 replicates
basepairs
```{r}
all_vs_all_overlap %>%
  filter( str_detect(set1, "P7") &
           str_detect(set2, "P7") )  %>%
  select(!starts_with("union")) %>%
  plot_overlaps()

```
elements
```{r}
all_vs_all_overlap_by_element %>%
  filter( str_detect(set1, "P7") &
           str_detect(set2, "P7") )  %>%
  select(!starts_with("union")) %>%
  plot_overlaps()

```
### Chip P60 replicates
```{r}
all_vs_all_overlap %>%
  filter( str_detect(set1, "P60") &
           str_detect(set2, "P60") )  %>%
  select(!starts_with("union")) %>%
  plot_overlaps()

```
By element
```{r}
all_vs_all_overlap_by_element %>%
  filter( str_detect(set1, "P60") &
           str_detect(set2, "P60") )  %>%
  select(!starts_with("union")) %>%
  plot_overlaps()

```


Notes 

Overlap of ChIP and CutNRun
- I think the disparity is because ChIP peaks are larger than CutNRun
- Because ChIP peaks are wider than CutNRun peaks, we expect that for peaks that are recovered in both assays, most bps in ChIP are not in CutNRun
- Doing overlaps by element addresses this issue

Within reps CutNRun
- Both overlaps by basepairs and by element give similar results because when comparing replicates, both sets of regions are of similar size and widths. 
- However, because regions are small, there might be a case when a single "real peak" is detected in two different experiments, but the reported regions are adjacent if not overlapping.
- Alternatively, there might be more experimental noise:
  - looking at other CutNRun datasets would help
  - Using more stringent cutoff values, we would expect a higher overlap

Next steps 
- Try merging Cut-N-Run peaks if they are close
- Try phyper or bedtools fisher for overlaps
- Check https://bioconductor.org/packages/release/bioc/vignettes/ChIPpeakAnno/inst/doc/pipeline.html

