---
title: "zic_cutnrun_chip_reprocessed_overlaps"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

How similar are in vitro and in vivo Zic peaks?

- CutNRun overlap with ChIP peaks
  - Load CutNrun and Chip peaks:
    - CutNrun:
      - seacr peaks
      - macs peaks
    - ChIP:
      - seacr peaks
      - macs peaks
      - previous macs peaks
  - Overlap all vs all
  - Plot overlaps:
    - CutNrun:
      - seacr between replicates
      - macs between replicates
      - seacr vs macs
    - ChIP:
      - seacr between replicates
      - macs between replicates
      - seacr vs macs
      - new vs old macs
    - CutNrun vs Chip:
      - seacr
      - macs
      

# Load CutNrun and Chip peaks

## load cutnrun peaks as granges objects seacr
```{r}
library(rtracklayer)
library(tidyverse)

cutnrun_seacr_dir <- "../data_output/cutnrun_zic/seacr_peaks/"
cutnrun_seacr_files <- list.files(cutnrun_seacr_dir,pattern = ".bed")
cutnrun_seacr_full <- paste0(cutnrun_seacr_dir,cutnrun_seacr_files)


cutnrun_seacr_granges <- lapply(cutnrun_seacr_full, import.bedGraph)

names(cutnrun_seacr_granges) <- str_split(cutnrun_seacr_files,"_",n=2) %>% map_chr(1)

names(cutnrun_seacr_granges) <- paste0("cutnrun-seacr-",names(cutnrun_seacr_granges))

```

macs
```{r}

cutnrun_macs_dir <- "../data_output/cutnrun_zic/macs_peaks/"
cutnrun_macs_files <- list.files(cutnrun_macs_dir, pattern = "narrowPeak$")
cutnrun_macs_full <- paste0(cutnrun_macs_dir,cutnrun_macs_files)

cutnrun_macs_granges <- lapply(cutnrun_macs_full,
                               import,
                               format = "BED",
                               extraCols = c(singnalValue = "numeric",
                                             pValue = "numeric",
                                             qValue = "numeric",
                                             peak = "integer"))

names(cutnrun_macs_granges) <- str_split(cutnrun_macs_files,"_",n=2) %>% map_chr(1)

names(cutnrun_macs_granges) <- paste0("cutnrun-macs-",names(cutnrun_macs_granges))

```


## load ChIP peaks - previous macs
```{r}
chip_peaks_dir <- "../data_output/bed/"
chip_peaks_files <- list.files(chip_peaks_dir, pattern = "mm10.bed")
chip_peaks_full <- paste0(chip_peaks_dir,chip_peaks_files)
chip_names <- str_split(chip_peaks_files,"_",n=3) %>% map_chr(2)
chip_names <- paste0("ChIP-macs1-","rep",rep(1:2,2),"-",chip_names)

chip_macs1_granges <- lapply(chip_peaks_full, import.bed)

names(chip_macs1_granges) <- chip_names

```

## load ChIP peaks - reprocessed macs
```{r}
chip_macs2_peaks_dir <- "../data_output/chip_zic/chip_zic_macs2_peaks/"
chip_macs2_peaks_files <- list.files(chip_macs2_peaks_dir)
chip_macs2_peaks_files <- chip_macs2_peaks_files[1:4]
chip_macs2_peaks_full <- paste0(chip_macs2_peaks_dir,chip_macs2_peaks_files)
chip_macs2_peaks_names <- c("P7","P7","P60","P60")

chip_macs2_peaks_names <- paste0("ChIP-macs2-","rep",rep(1:2,2),"-",chip_macs2_peaks_names)


chip_macs2_granges <- lapply(chip_macs2_peaks_full,
                               import,
                               format = "BED",
                               extraCols = c(singnalValue = "numeric",
                                             pValue = "numeric",
                                             qValue = "numeric",
                                             peak = "integer"))


names(chip_macs2_granges) <- chip_macs2_peaks_names

```

## load ChIP peaks - reprocessed seacr
```{r}
chip_seacr_peaks_dir <- "../data_output/chip_zic/chip_zic_seacr_peaks/"
chip_seacr_peaks_files <- list.files(chip_seacr_peaks_dir,pattern = "bed$")
chip_seacr_peaks_files <- chip_seacr_peaks_files[1:4]
chip_seacr_peaks_full <- paste0(chip_seacr_peaks_dir,chip_seacr_peaks_files)
chip_seacr_peaks_names <- c("P7","P7","P60","P60")
chip_seacr_peaks_names <- paste0("ChIP-seacr-","rep",rep(1:2,2),"-",chip_seacr_peaks_names)

chip_seacr_granges <- lapply(chip_seacr_peaks_full, import.bedGraph)

names(chip_seacr_granges) <- chip_seacr_peaks_names

```

# Overlap all vs all

```{r}
library(GenomicRanges)

all_peaks <- GRangesList(c(cutnrun_macs_granges,
                           cutnrun_seacr_granges,
                           chip_macs1_granges,
                           chip_macs2_granges,
                           chip_seacr_granges))

```

Remove peaks that overlap blacklist
```{r}
mm10_blacklist <- import("../data_input/bed/mm10-blacklist.v2.bed.gz")

all_peaks <- lapply(all_peaks, function(x){
  x[!overlapsAny(x,mm10_blacklist)]
})

```


In basepairs

```{r}

all_vs_all_overlap_by_basepairs <- combn(1:length(all_peaks),2,
                            FUN = function(x) {
                              intersection <- sum(width(GenomicRanges::intersect(all_peaks[[x[1]]],all_peaks[[x[2]]])))
                              setdiff1 <- sum(width(GenomicRanges::setdiff(all_peaks[[x[1]]],all_peaks[[x[2]]])))
                              setdiff2 <- sum(width(GenomicRanges::setdiff(all_peaks[[x[2]]],all_peaks[[x[1]]])))
                              union <- sum(width(GenomicRanges::union(all_peaks[[x[1]]],all_peaks[[x[2]]])))
                              return(data.frame("intersection" = intersection,
                                                "union" = union,
                                                "set1" = names(all_peaks[x[1]]),
                                                "set2" = names(all_peaks[x[2]]),
                                                "setdiff1" = setdiff1,
                                                "setdiff2" = setdiff2)
                              )
                            },
                            simplify = FALSE)

all_vs_all_overlap_by_basepairs <- do.call(rbind,all_vs_all_overlap_by_basepairs)


```

```{r}
get_overlaps_by_basepair <- function(peaks_list) {
  peaks_list_olap <- combn(1:length(peaks_list),2,
                           FUN = function(x) {
                             intersection <- sum(width(GenomicRanges::intersect(peaks_list[[x[1]]],peaks_list[[x[2]]])))
                             setdiff1 <- sum(width(GenomicRanges::setdiff(peaks_list[[x[1]]],peaks_list[[x[2]]])))
                             setdiff2 <- sum(width(GenomicRanges::setdiff(peaks_list[[x[2]]],peaks_list[[x[1]]])))
                             union <- sum(width(GenomicRanges::union(peaks_list[[x[1]]],peaks_list[[x[2]]])))
                             return(data.frame("intersection" = intersection,
                                               "union" = union,
                                               "set1" = names(peaks_list[x[1]]),
                                               "set2" = names(peaks_list[x[2]]),
                                               "setdiff1" = setdiff1,
                                               "setdiff2" = setdiff2)
                             )
                           },
                           simplify = FALSE)
  
  peaks_list_olap <- do.call(rbind,peaks_list_olap)
  return(peaks_list_olap)
}


```




By element

```{r}
get_overlaps_by_element <- function(peaks_list) {
  peaks_list_olap <- combn(1:length(peaks_list),2,
                            FUN = function(x) {
                              intersection <- sum(IRanges::overlapsAny(peaks_list[[x[1]]],peaks_list[[x[2]]]))
                              setdiff1 <- length(peaks_list[[x[1]]]) - intersection
                              setdiff2 <-length(peaks_list[[x[2]]]) - intersection 
                              
                              return(data.frame("intersection" = intersection,
                                                "set1" = names(peaks_list[x[1]]),
                                                "set2" = names(peaks_list[x[2]]),
                                                "setdiff1" = setdiff1,
                                                "setdiff2" = setdiff2)
                              )
                            },
                            simplify = FALSE)

peaks_list_olap <- do.call(rbind,peaks_list_olap)
return(peaks_list_olap)
}

get_overlaps_by_element2 <- function(peaks_list) {
  peaks_list_olap <- combn(1:length(peaks_list),2,
                            FUN = function(x) {
                              intersection1 <- sum(IRanges::overlapsAny(peaks_list[[x[1]]],peaks_list[[x[2]]]))
                              intersection2 <- sum(IRanges::overlapsAny(peaks_list[[x[2]]],peaks_list[[x[1]]]))

                              setdiff1 <- length(peaks_list[[x[1]]]) - intersection1
                              setdiff2 <-length(peaks_list[[x[2]]]) - intersection2 
                              
                              return(data.frame("intersection1" = intersection1,
                                                "intersection2" = intersection2,
                                                "set1" = names(peaks_list[x[1]]),
                                                "set2" = names(peaks_list[x[2]]),
                                                "setdiff1" = setdiff1,
                                                "setdiff2" = setdiff2)
                              )
                            },
                            simplify = FALSE)

peaks_list_olap <- do.call(rbind,peaks_list_olap)
return(peaks_list_olap)
}
```


```{r}

all_vs_all_overlap_by_element <- combn(1:length(all_peaks),2,
                            FUN = function(x) {
                              intersection <- sum(IRanges::overlapsAny(all_peaks[[x[1]]],all_peaks[[x[2]]]))
                              setdiff1 <- length(all_peaks[[x[1]]]) - intersection
                              setdiff2 <-length(all_peaks[[x[2]]]) - intersection 
                              
                              return(data.frame("intersection" = intersection,
                                                "set1" = names(all_peaks[x[1]]),
                                                "set2" = names(all_peaks[x[2]]),
                                                "setdiff1" = setdiff1,
                                                "setdiff2" = setdiff2)
                              )
                            },
                            simplify = FALSE)

all_vs_all_overlap_by_element <- do.call(rbind,all_vs_all_overlap_by_element)


```

# Plots

```{r}
plot_overlaps <- function(overlaps_df) {
  overlaps_df %>%
  pivot_longer(is.numeric) %>%
    group_by(set1,set2) %>%
    mutate(percentage =100 * value / sum(value)) %>%
    ungroup() %>%
    mutate(name = fct_relevel(name, "setdiff2","setdiff1","intersection"),
           comparison = factor(paste(set1,set2))) %>%
    ggplot(aes(y = comparison, x = percentage, fill = name)) +
    geom_col(width = 0.5) +
    ylab(NULL) +
    xlab("Percentage Overlap") +
    scale_fill_viridis_d(begin = 0.2,end = 0.8) +
    theme_minimal() 
}
```


# Plot overlaps
## CutNrun

### seacr vs macs between replicates

#### Day 1

By basepair
```{r}
library(patchwork)

(all_vs_all_overlap_by_basepairs %>%
  filter(str_detect(set1, "macs") &
           str_detect(set2, "macs")) %>%
  filter( str_detect(set1, "D1") &
           str_detect(set2, "D1") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)") +
  ggtitle("macs")) /
  (all_vs_all_overlap_by_basepairs %>%
  filter(str_detect(set1, "seacr") &
           str_detect(set2, "seacr")) %>%
  filter( str_detect(set1, "D1") &
           str_detect(set2, "D1") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)") +
  ggtitle("seacr")) 
```

By element
```{r}
(all_vs_all_overlap_by_element %>%
  filter(str_detect(set1, "macs") &
           str_detect(set2, "macs")) %>%
  filter( str_detect(set1, "D1") &
           str_detect(set2, "D1") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)") +
  ggtitle("macs")) /
  (all_vs_all_overlap_by_element %>%
  filter(str_detect(set1, "seacr") &
           str_detect(set2, "seacr")) %>%
  filter( str_detect(set1, "D1") &
           str_detect(set2, "D1") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)") +
  ggtitle("seacr")) 
```

#### Day 3

By basepair
```{r}


(all_vs_all_overlap_by_basepairs %>%
  filter(str_detect(set1, "macs") &
           str_detect(set2, "macs")) %>%
  filter( str_detect(set1, "D3") &
           str_detect(set2, "D3") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)") +
  ggtitle("macs")) /
  (all_vs_all_overlap_by_basepairs %>%
  filter(str_detect(set1, "seacr") &
           str_detect(set2, "seacr")) %>%
  filter( str_detect(set1, "D3") &
           str_detect(set2, "D3") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)") +
  ggtitle("seacr")) 
```

By element
```{r}
(all_vs_all_overlap_by_element %>%
  filter(str_detect(set1, "macs") &
           str_detect(set2, "macs")) %>%
  filter( str_detect(set1, "D3") &
           str_detect(set2, "D3") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)") +
  ggtitle("macs")) /
  (all_vs_all_overlap_by_element %>%
  filter(str_detect(set1, "seacr") &
           str_detect(set2, "seacr")) %>%
  filter( str_detect(set1, "D3") &
           str_detect(set2, "D3") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)") +
  ggtitle("seacr")) 
```

#### Day 5

By basepair
```{r}
library(patchwork)

(all_vs_all_overlap_by_basepairs %>%
  filter(str_detect(set1, "macs") &
           str_detect(set2, "macs")) %>%
  filter( str_detect(set1, "D5") &
           str_detect(set2, "D5") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)") +
  ggtitle("macs")) /
  (all_vs_all_overlap_by_basepairs %>%
  filter(str_detect(set1, "seacr") &
           str_detect(set2, "seacr")) %>%
  filter( str_detect(set1, "D5") &
           str_detect(set2, "D5") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)") +
  ggtitle("seacr")) 
```

By element
```{r}
(all_vs_all_overlap_by_element %>%
  filter(str_detect(set1, "macs") &
           str_detect(set2, "macs")) %>%
  filter( str_detect(set1, "D5") &
           str_detect(set2, "D5") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)") +
  ggtitle("macs")) /
  (all_vs_all_overlap_by_element %>%
  filter(str_detect(set1, "seacr") &
           str_detect(set2, "seacr")) %>%
  filter( str_detect(set1, "D5") &
           str_detect(set2, "D5") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)") +
  ggtitle("seacr")) 
```

#### Day 7

By basepair
```{r}


(all_vs_all_overlap_by_basepairs %>%
  filter(str_detect(set1, "macs") &
           str_detect(set2, "macs")) %>%
  filter( str_detect(set1, "D7") &
           str_detect(set2, "D7") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)") +
  ggtitle("macs")) /
  (all_vs_all_overlap_by_basepairs %>%
  filter(str_detect(set1, "seacr") &
           str_detect(set2, "seacr")) %>%
  filter( str_detect(set1, "D7") &
           str_detect(set2, "D7") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)") +
  ggtitle("seacr")) 
```

By element
```{r}
(all_vs_all_overlap_by_element %>%
  filter(str_detect(set1, "macs") &
           str_detect(set2, "macs")) %>%
  filter( str_detect(set1, "D7") &
           str_detect(set2, "D7") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)") +
  ggtitle("macs")) /
  (all_vs_all_overlap_by_element %>%
  filter(str_detect(set1, "seacr") &
           str_detect(set2, "seacr")) %>%
  filter( str_detect(set1, "D7") &
           str_detect(set2, "D7") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)") +
  ggtitle("seacr")) 
```



### seacr vs macs

By basepair
```{r}

all_vs_all_overlap_by_basepairs %>%
  separate(set1, into =c("exp1","caller1","rep1","time1") ,remove = FALSE) %>%
  separate(set2, into =c("exp2","caller2","rep2","time2") , remove = FALSE) %>%
  filter(exp1 == "cutnrun",
         exp2 == "cutnrun",
         rep1 == rep2,
         time1 == time2) %>%
  select(-union) %>%
  plot_overlaps() +
  xlab("Overlap (BP)") 

```

By element
```{r}

all_vs_all_overlap_by_element %>%
  separate(set1, into =c("exp1","caller1","rep1","time1") ,remove = FALSE) %>%
  separate(set2, into =c("exp2","caller2","rep2","time2") , remove = FALSE) %>%
  filter(exp1 == "cutnrun",
         exp2 == "cutnrun",
         rep1 == rep2,
         time1 == time2) %>%
  plot_overlaps() +
  xlab("Overlap (N)") 

```

## ChIP:
### seacr and macs between replicates

#### Day P7

By basepair
```{r}

(all_vs_all_overlap_by_basepairs %>%
  filter(str_detect(set1, "macs2") &
           str_detect(set2, "macs2")) %>%
  filter( str_detect(set1, "P7") &
           str_detect(set2, "P7") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)") +
  ggtitle("macs")) /
  (all_vs_all_overlap_by_basepairs %>%
  filter(str_detect(set1, "seacr") &
           str_detect(set2, "seacr")) %>%
  filter( str_detect(set1, "P7") &
           str_detect(set2, "P7") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)") +
  ggtitle("seacr")) 
```

By element
```{r}

(all_vs_all_overlap_by_element %>%
  filter(str_detect(set1, "macs2") &
           str_detect(set2, "macs2")) %>%
  filter( str_detect(set1, "P7") &
           str_detect(set2, "P7") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)") +
  ggtitle("macs")) /
  (all_vs_all_overlap_by_element %>%
  filter(str_detect(set1, "seacr") &
           str_detect(set2, "seacr")) %>%
  filter( str_detect(set1, "P7") &
           str_detect(set2, "P7") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)") +
  ggtitle("seacr")) 
```

#### Day P60

By basepair
```{r}

(all_vs_all_overlap_by_basepairs %>%
  filter(str_detect(set1, "macs2") &
           str_detect(set2, "macs2")) %>%
  filter( str_detect(set1, "P60") &
           str_detect(set2, "P60") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)") +
  ggtitle("macs")) /
  (all_vs_all_overlap_by_basepairs %>%
  filter(str_detect(set1, "seacr") &
           str_detect(set2, "seacr")) %>%
  filter( str_detect(set1, "P60") &
           str_detect(set2, "P60") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)") +
  ggtitle("seacr")) 
```

By element
```{r}

(all_vs_all_overlap_by_element %>%
  filter(str_detect(set1, "macs2") &
           str_detect(set2, "macs2")) %>%
  filter( str_detect(set1, "P60") &
           str_detect(set2, "P60") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)") +
  ggtitle("macs")) /
  (all_vs_all_overlap_by_element %>%
  filter(str_detect(set1, "seacr") &
           str_detect(set2, "seacr")) %>%
  filter( str_detect(set1, "P60") &
           str_detect(set2, "P60") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)") +
  ggtitle("seacr")) 
```

### seacr vs macs

By basepair
```{r}

all_vs_all_overlap_by_basepairs %>%
  separate(set1, into =c("exp1","caller1","rep1","time1") ,remove = FALSE) %>%
  separate(set2, into =c("exp2","caller2","rep2","time2") , remove = FALSE) %>%
  filter(exp1 == "ChIP",
         exp2 == "ChIP",
         rep1 == rep2,
         time1 == time2) %>%
  select(-union) %>%
  plot_overlaps() +
  xlab("Overlap (BP)") 

```

By element
```{r}

all_vs_all_overlap_by_element %>%
  separate(set1, into =c("exp1","caller1","rep1","time1") ,remove = FALSE) %>%
  separate(set2, into =c("exp2","caller2","rep2","time2") , remove = FALSE) %>%
  filter(exp1 == "ChIP",
         exp2 == "ChIP",
         rep1 == rep2,
         time1 == time2) %>%
  plot_overlaps() +
  xlab("Overlap (N)") 

```


### new vs old macs

#### Day P7

By basepair
```{r}

(all_vs_all_overlap_by_basepairs %>%
  filter(str_detect(set1, "macs2") &
           str_detect(set2, "macs2")) %>%
  filter( str_detect(set1, "P7") &
           str_detect(set2, "P7") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)") +
  ggtitle("macs2")) /
  (all_vs_all_overlap_by_basepairs %>%
  filter(str_detect(set1, "macs1") &
           str_detect(set2, "macs1")) %>%
  filter( str_detect(set1, "P7") &
           str_detect(set2, "P7") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)") +
  ggtitle("macs1")) 
```

By element
```{r}

(all_vs_all_overlap_by_element %>%
  filter(str_detect(set1, "macs2") &
           str_detect(set2, "macs2")) %>%
  filter( str_detect(set1, "P7") &
           str_detect(set2, "P7") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)") +
  ggtitle("macs2")) /
  (all_vs_all_overlap_by_element %>%
  filter(str_detect(set1, "macs1") &
           str_detect(set2, "macs1")) %>%
  filter( str_detect(set1, "P7") &
           str_detect(set2, "P7") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)") +
  ggtitle("macs1")) 
```

#### Day P60

By basepair
```{r}

(all_vs_all_overlap_by_basepairs %>%
  filter(str_detect(set1, "macs2") &
           str_detect(set2, "macs2")) %>%
  filter( str_detect(set1, "P60") &
           str_detect(set2, "P60") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)") +
  ggtitle("macs2")) /
  (all_vs_all_overlap_by_basepairs %>%
  filter(str_detect(set1, "macs1") &
           str_detect(set2, "macs1")) %>%
  filter( str_detect(set1, "P60") &
           str_detect(set2, "P60") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)") +
  ggtitle("macs1")) 
```

By element
```{r}

(all_vs_all_overlap_by_element %>%
  filter(str_detect(set1, "macs2") &
           str_detect(set2, "macs2")) %>%
  filter( str_detect(set1, "P60") &
           str_detect(set2, "P60") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)") +
  ggtitle("macs2")) /
  (all_vs_all_overlap_by_element %>%
  filter(str_detect(set1, "macs1") &
           str_detect(set2, "macs1")) %>%
  filter( str_detect(set1, "P60") &
           str_detect(set2, "P60") ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)") +
  ggtitle("macs1")) 
```


## CutNrun vs Chip:

### seacr

#### Matched times plots

##### Culture day 1 vs in vivo day 7

Basepairs
```{r}
all_vs_all_overlap_by_basepairs %>%
  filter(str_detect(set1, "seacr"),
         str_detect(set2, "seacr")) %>%
  filter( (str_detect(set1, "cutnrun") &
           str_detect(set2, "ChIP") ) | (
             str_detect(set1, "ChIP") &
           str_detect(set2, "cutnrun") )
           ) %>%
  filter( (str_detect(set1, "D1") &
           str_detect(set2, "P7") ) | (
             str_detect(set1, "P7") &
           str_detect(set2, "D1") )
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)")

```

By element
```{r}
all_vs_all_overlap_by_element %>%
  filter(str_detect(set1, "seacr"),
         str_detect(set2, "seacr")) %>%
  filter( (str_detect(set1, "cutnrun") &
           str_detect(set2, "ChIP") ) | (
             str_detect(set1, "ChIP") &
           str_detect(set2, "cutnrun") )
           ) %>%
  filter( (str_detect(set1, "D1") &
           str_detect(set2, "P7") ) | (
             str_detect(set1, "P7") &
           str_detect(set2, "D1") )
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)")

```


##### Culture day 3 vs in vivo day 7
Basepairs
```{r}
all_vs_all_overlap_by_basepairs %>%
  filter(str_detect(set1, "seacr"),
         str_detect(set2, "seacr")) %>%
  filter( (str_detect(set1, "cutnrun") &
           str_detect(set2, "ChIP") ) | (
             str_detect(set1, "ChIP") &
           str_detect(set2, "cutnrun") )
           ) %>%
  filter( (str_detect(set1, "D3") &
           str_detect(set2, "P7") ) | (
             str_detect(set1, "P7") &
           str_detect(set2, "D3") )
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)")

```

By element
```{r}
all_vs_all_overlap_by_element %>%
  filter(str_detect(set1, "seacr"),
         str_detect(set2, "seacr")) %>%
  filter( (str_detect(set1, "cutnrun") &
           str_detect(set2, "ChIP") ) | (
             str_detect(set1, "ChIP") &
           str_detect(set2, "cutnrun") )
           ) %>%
  filter( (str_detect(set1, "D3") &
           str_detect(set2, "P7") ) | (
             str_detect(set1, "P7") &
           str_detect(set2, "D3") )
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)")

```
##### Culture day 7 vs in vivo day 60

Basepairs
```{r}
all_vs_all_overlap_by_basepairs %>%
  filter(str_detect(set1, "seacr"),
         str_detect(set2, "seacr")) %>%
  filter( (str_detect(set1, "cutnrun") &
           str_detect(set2, "ChIP") ) | (
             str_detect(set1, "ChIP") &
           str_detect(set2, "cutnrun") )
           ) %>%
  filter( (str_detect(set1, "D7") &
           str_detect(set2, "P60") ) | (
             str_detect(set1, "P60") &
           str_detect(set2, "D7") )
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)")

```

By element
```{r}
all_vs_all_overlap_by_element %>%
  filter(str_detect(set1, "seacr"),
         str_detect(set2, "seacr")) %>%
  filter( (str_detect(set1, "cutnrun") &
           str_detect(set2, "ChIP") ) | (
             str_detect(set1, "ChIP") &
           str_detect(set2, "cutnrun") )
           ) %>%
  filter( (str_detect(set1, "D7") &
           str_detect(set2, "P60") ) | (
             str_detect(set1, "P60") &
           str_detect(set2, "D7") )
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)")

```
### macs

#### Matched times plots

##### Culture day 1 vs in vivo day 7

Basepairs
```{r}
all_vs_all_overlap_by_basepairs %>%
  filter( (str_detect(set1, "cutnrun-macs") &
           str_detect(set2, "ChIP-macs2") ) | (
             str_detect(set1, "ChIP-macs2") &
           str_detect(set2, "cutnrun-macs") )
           ) %>%
  filter( (str_detect(set1, "D1") &
           str_detect(set2, "P7") ) | (
             str_detect(set1, "P7") &
           str_detect(set2, "D1") )
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)")

```

By element
```{r}
all_vs_all_overlap_by_element %>%
  filter( (str_detect(set1, "cutnrun-macs") &
           str_detect(set2, "ChIP-macs2") ) | (
             str_detect(set1, "ChIP-macs2") &
           str_detect(set2, "cutnrun-macs") )
           ) %>%
  filter( (str_detect(set1, "D1") &
           str_detect(set2, "P7") ) | (
             str_detect(set1, "P7") &
           str_detect(set2, "D1") )
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)")

```


##### Culture day 3 vs in vivo day 7
Basepairs
```{r}
all_vs_all_overlap_by_basepairs %>%
  filter( (str_detect(set1, "cutnrun-macs") &
           str_detect(set2, "ChIP-macs2") ) | (
             str_detect(set1, "ChIP-macs2") &
           str_detect(set2, "cutnrun-macs") )
           ) %>%
  filter( (str_detect(set1, "D3") &
           str_detect(set2, "P7") ) | (
             str_detect(set1, "P7") &
           str_detect(set2, "D3") )
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)")

```

By element
```{r}
all_vs_all_overlap_by_element %>%
  filter( (str_detect(set1, "cutnrun-macs") &
           str_detect(set2, "ChIP-macs2") ) | (
             str_detect(set1, "ChIP-macs2") &
           str_detect(set2, "cutnrun-macs") )
           ) %>%
  filter( (str_detect(set1, "D3") &
           str_detect(set2, "P7") ) | (
             str_detect(set1, "P7") &
           str_detect(set2, "D3") )
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)")

```
##### Culture day 7 vs in vivo day 60

Basepairs
```{r}
all_vs_all_overlap_by_basepairs %>%
  filter( (str_detect(set1, "cutnrun-macs") &
           str_detect(set2, "ChIP-macs2") ) | (
             str_detect(set1, "ChIP-macs2") &
           str_detect(set2, "cutnrun-macs") )
           ) %>%
  filter( (str_detect(set1, "D7") &
           str_detect(set2, "P60") ) | (
             str_detect(set1, "P60") &
           str_detect(set2, "D7") )
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)")

```

By element
```{r}
all_vs_all_overlap_by_element %>%
  filter( (str_detect(set1, "cutnrun-macs") &
           str_detect(set2, "ChIP-macs2") ) | (
             str_detect(set1, "ChIP-macs2") &
           str_detect(set2, "cutnrun-macs") )
           ) %>%
  filter( (str_detect(set1, "D7") &
           str_detect(set2, "P60") ) | (
             str_detect(set1, "P60") &
           str_detect(set2, "D7") )
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)")

```


To more closely answer the experimental question of how similar is Zic binding in vivo vs in vitro, we could take the most reproducible peaks from each experiment before doing the between experiment comparisons.


0. Get macs2 peaks
```{r}
macs2_peaks <- all_peaks[str_detect(names(all_peaks),"cutnrun-macs|ChIP-macs2")]

macs2_peaks_list <- split(macs2_peaks, str_split(names(macs2_peaks),"-",4) %>% map_chr(4))

```

1. Get intersection of peaks for all replicates within each timepoint 
```{r}
macs2_peaks_reduced_list <- lapply(macs2_peaks_list, function(x) {
  Reduce(GenomicRanges::intersect, x)
})

```

2. Get All vs all overlaps
```{r}


macs2_peaks_reduced_overlap_by_element <- get_overlaps_by_element(macs2_peaks_reduced_list)

macs2_peaks_reduced_overlap_by_basepair <- get_overlaps_by_basepair(macs2_peaks_reduced_list)

macs2_peaks_reduced_overlap_by_element2 <- get_overlaps_by_element2(macs2_peaks_reduced_list)

```

3. Plot overlaps

D1 vs P7
```{r}
macs2_peaks_reduced_overlap_by_element %>%
    filter( (str_detect(set1, "P") &
           str_detect(set2, "D") ) | (
             str_detect(set1, "D") &
           str_detect(set2, "P") )
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)")
```

```{r}
macs2_peaks_reduced_overlap_by_element2 %>%
    filter( (str_detect(set1, "P") &
           str_detect(set2, "D") ) | (
             str_detect(set1, "D") &
           str_detect(set2, "P") )
           ) %>%
  select(!starts_with("union"),-intersection2,-setdiff2) %>%
  plot_overlaps() +
  xlab("Overlap (N)")

macs2_peaks_reduced_overlap_by_element2 %>%
    filter( (str_detect(set1, "P") &
           str_detect(set2, "D") ) | (
             str_detect(set1, "D") &
           str_detect(set2, "P") )
           ) %>%
  select(!starts_with("union"),-intersection1,-setdiff1) %>%
  plot_overlaps() +
  xlab("Overlap (N)")

```


```{r}
macs2_peaks_reduced_overlap_by_element %>%
    filter( (str_detect(set1, "P") &
           str_detect(set2, "D") ) | (
             str_detect(set1, "D") &
           str_detect(set2, "P") )
           ) %>%
  select(!starts_with("union"), -setdiff2) %>%
  plot_overlaps() +
  xlab("Overlap (N)")
```


D1 vs P7
```{r}
macs2_peaks_reduced_overlap_by_basepair %>%
    filter( (str_detect(set1, "P") &
           str_detect(set2, "D") ) | (
             str_detect(set1, "D") &
           str_detect(set2, "P") )
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)")
```

D1 vs P7
```{r}
macs2_peaks_reduced_overlap_by_element %>%
    filter( (str_detect(set1, "D") &
           str_detect(set2, "D") ) 
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)")
```
```{r}
macs2_peaks_reduced_overlap_by_element %>%
    filter( (str_detect(set1, "P") &
           str_detect(set2, "P") ) 
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)")

```

Try with seacr
0. Get seacr peaks
```{r}
seacr_peaks <- all_peaks[str_detect(names(all_peaks),"seacr")]

seacr_peaks_list <- split(seacr_peaks, str_split(names(seacr_peaks),"-",4) %>% map_chr(4))

```

1. Get intersection of peaks for all replicates within each timepoint 
```{r}
seacr_peaks_reduced_list <- lapply(seacr_peaks_list, function(x) {
  Reduce(GenomicRanges::intersect, x)
})

```

2. Get All vs all overlaps
```{r}


seacr_peaks_reduced_overlap_by_element <- get_overlaps_by_element(seacr_peaks_reduced_list)

seacr_peaks_reduced_overlap_by_basepair <- get_overlaps_by_basepair(seacr_peaks_reduced_list)


seacr_peaks_reduced_overlap_by_element2 <- get_overlaps_by_element2(seacr_peaks_reduced_list)

```

3. Plot overlaps

D1 vs P7
```{r}
seacr_peaks_reduced_overlap_by_element %>%
    filter( (str_detect(set1, "P") &
           str_detect(set2, "D") ) | (
             str_detect(set1, "D") &
           str_detect(set2, "P") )
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)")
```
```{r}
seacr_peaks_reduced_overlap_by_basepair %>%
    filter( (str_detect(set1, "P") &
           str_detect(set2, "D") ) | (
             str_detect(set1, "D") &
           str_detect(set2, "P") )
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (BP)")
```

```{r}
seacr_peaks_reduced_overlap_by_element2 %>%
    filter( (str_detect(set1, "P") &
           str_detect(set2, "D") ) | (
             str_detect(set1, "D") &
           str_detect(set2, "P") )
           ) %>%
  select(!starts_with("union"),-intersection2,-setdiff2) %>%
  plot_overlaps() +
  xlab("Overlap (N)")

seacr_peaks_reduced_overlap_by_element2 %>%
    filter( (str_detect(set1, "P") &
           str_detect(set2, "D") ) | (
             str_detect(set1, "D") &
           str_detect(set2, "P") )
           ) %>%
  select(!starts_with("union"),-intersection1,-setdiff1) %>%
  plot_overlaps() +
  xlab("Overlap (N)")

```
D1 vs P7
```{r}
seacr_peaks_reduced_overlap_by_element %>%
    filter( (str_detect(set1, "D") &
           str_detect(set2, "D") ) 
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)")
```

```{r}
seacr_peaks_reduced_overlap_by_element %>%
    filter( (str_detect(set1, "P") &
           str_detect(set2, "P") ) 
           ) %>%
  select(!starts_with("union")) %>%
  plot_overlaps() +
  xlab("Overlap (N)")

```


Notes:

From looking at the browser
- New macs2 chip peaks look better than previous ones on the browser. They are shorter, and more centered on the signal summit.
- seacr chip peaks don't look as good as macs2 peaks. 
- Macs2 seems to work well for cutnrun as well. Calls less peaks where signal looks more noisy compared to seacr, but also misses some peaks that look real that seacr doesn't.

From looking at the overlaps
- When comparing macs2 vs seacr, most macs2 peaks are also called by seacr, but seacr calls many more. These are likely a mix of seacr false positives, and macs2 false negatives.
- If we want to compare cutnrun vs chip with the same caller, I would trust more the macs2 comparison, which works reasonably well on both datasets.

From looking at the stringent overlaps analysis
- In vitro Zic binding is progressively lost as days pass


